Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2019-12-06 manuscript/climate_sensitivity_draft.Rmd
Update climate_sensitivity_draft.Rmd
working on intro, Lawren's comments

- Individual-level drought resistance decreased with tree height, which was the dominant size-related variable affecting drought response. Resistance was greater both among species whose leaves lost turgor (wilted) at more negative water potentials, and whose leaves experienced less shrinkage upon dessication. However, there was substantial variation in the best predictor variables across the three drought periods.

- We conclude that hydraulic traits and tree height influence growth responses during drought, and can explain variation in the tree ring record spanning historical droughts. Thus, these traits can be useful for predicting future droughts under climate change.

*Key words*:  canopy position; drought; Forest Global Earth Observatory (ForestGEO); leaf hydraulic traits; temperate broadleaf deciduous forest; annual growth; tree height; tree-ring [5-8] *Five to eight key words (in alphabetical order) . Words that are in the title can, and should, be among these.*

\newpage
### Introduction
**See comment 4,31**
Forests globally play a critical role in climate regulation [@bonan_forests_2008], yet there remains enormous uncertainty as to how the terrestrial carbon (C) sink, which is dominated by forests, will respond to climate change [@friedlingstein_climatecarbon_2006]. An important aspect of this uncertainty lies with physiological responses of trees to drought [@kennedy_implementing_2019]. In many forested regions around the world, the risk of severe drought is increasing [@trenberth_global_2014; @cook_global_2014; @cook_unprecedented_2015; @mankin_mid-latitude_2019], sometimes despite increasing precipitation [@intergovernmental_panel_on_climate_change_climate_2015]. Droughts resulting from global change have been affecting forests worldwide [@allen_global_2010], and are expected to continue severely impacting forests [@anderegg_consequences_2013; @adams_temperature_2009; @breda_temperate_2006; @hartmann_research_2018; @allen_underestimation_2015; @abrams_interdisciplinary_2016; @adams_multi-species_2017]. During drought, larger trees generally suffer greater reductions in growth and increased mortality rates [e.g., @bennett_larger_2015; @stovall_tree_2019], resulting in disproportionate impacts on forest C storage [@meakem_role_2018]. As a result, forest drought responses stand to strongly impact forest feedbacks to climate change [@mcdowell_multi-scale_2016; @charney_observed_2016]. Yet, accurate characterization of drought responses remains a modeling challenge [@kennedy_implementing_2019]-- in part because some of the underlying mechanisms remain unclear (**REF**). Understanding forest responses to drought requires increased functional understanding of how tree size, microenvironment, and species' traits jointly confer individual-level vulnerability or resistance, and the extent to which their influence is consistent across droughts.

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought. @bennett_larger_2015 was the first to show larger growth reductions for larger trees on a global scale, and numerous subsequent studies have reinforced this finding (e.g., @stovall_tree_2019; @hacket-pain_consistent_2016]. However, they quantified tree size based on DBH, which has no direct mechanistic response. Their study proposed two major mechanisms--besides the tendency for bark beetles to preferentially attack larger trees [@pfeifer_observations_2011]--for the observed greater drought growth reductions of large trees. First, @bennett_larger_2015 found taller trees have a greater biophysical challenge of lifting water greater distances against the effects of gravity and friction [@mcdowell_relationships_2011; @mcdowell_darcys_2015; @ryan_hydraulic_2006; @couvreur_water_2018], especially during drought [@zhang_size-dependent_2009]. Second, larger trees may have lower drought resistance because of their higher crown positions, where they are exposed to higher solar radiation, greater wind speeds, lower humidity, and lower CO_2_ concentrations [@koike_leaf_2001; @kunert_revised_2017 **REFS-KAT**). Subcanopy trees subsequently tend to fare better specifically due to the benefits of a buffered environment [@pretzsch_drought_2018]. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems confer a potential advantage in terms of allowing greater access to water; however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. In addition to what was described by @bennett_larger_2015, a final mechanism that could mediate tree size-related responses to drought is how hydraulic traits are distributed with respect to size [@meakem_role_2018; @liu_hydraulic_2019 [Roskilly et al. 2019](https://www.pnas.org/content/pnas/early/2019/06/14/1900734116.full.pdf)]. The pattern observed by @bennett_larger_2015 could be caused if the larger size classes were dominated by species less adapted to handle drought. Alternatively, larger size classes may be dominated by species that are inherently better adapted to greater biophysical challenges--as is the case in tropical moist forests of Panama, where larger size classes contain greater proportions of deciduous species [@condit_quantifying_2000; @meakem_role_2018]. Understanding the mechanisms driving more severe suffering of larger trees during drought will require sorting out the interactive effects of height, canopy position, soil water availability, and species' traits.

>>>>

- Individual-level drought resistance decreased with tree height, which was the dominant size-related variable affecting drought response. Resistance was greater  among species whose leaves lost turgor (wilted) at more negative water potentials, and whose leaves experienced less shrinkage upon dessication. However, there was substantial variation in the best predictor variables across the three drought periods.

- We conclude that hydraulic traits and tree height influence growth responses during drought, and can explain variation in the tree ring record spanning historical droughts. Thus, these traits can be useful for predicting future drought responses under climate change.

*Key words*: annual growth; canopy position; drought; Forest Global Earth Observatory (ForestGEO); leaf hydraulic traits; temperate broadleaf deciduous forest; tree height; tree-ring [5-8] *Five to eight key words (in alphabetical order) . Words that are in the title can, and should, be among these.*

\newpage
### Introduction
**See comment 4,31**
Forests globally play a critical role in climate regulation [@bonan_forests_2008], yet there remains enormous uncertainty as to how the terrestrial carbon (C) sink, which is dominated by forests, will respond to climate change [@friedlingstein_climatecarbon_2006]. An important aspect of this uncertainty lies with physiological responses of trees to drought [@kennedy_implementing_2019]. In many forested regions around the world, the risk of severe drought is increasing [@trenberth_global_2014; @cook_global_2014; @cook_unprecedented_2015; @mankin_mid-latitude_2019], sometimes despite increasing precipitation [@intergovernmental_panel_on_climate_change_climate_2015]. Droughts, exasperated by climate change, have been affecting forests worldwide [@allen_global_2010], and are expected to continue severely impacting forests [@anderegg_consequences_2013; @adams_temperature_2009; @breda_temperate_2006; @hartmann_research_2018; @allen_underestimation_2015; @abrams_interdisciplinary_2016; @adams_multi-species_2017]**too many refs here. reduce to 2-3**. However, the traits that predict tree growth and mortality responses to drought have been under debate, and are difficult to resolve with the limited available census data, which only rarely includes extreme drought. The aim of this study was to test how tree size and traits shaped growth responses to historical droughts in the tree-ring record.

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought. @bennett_larger_2015 was the first to show greater growth reductions for larger trees on a global scale, and numerous subsequent studies have reinforced this finding (*e.g.*, @stovall_tree_2019; @hacket-pain_consistent_2016]. However, it has yet to be resolved which of several potential underlying mechanisms drive this pattern. First, taller trees have a greater biophysical challenge of lifting water greater distances against the effects of gravity and friction [@mcdowell_relationships_2011; @mcdowell_darcys_2015; @ryan_hydraulic_2006; @couvreur_water_2018], especially during drought [@zhang_size-dependent_2009]. Second, larger trees may have lower drought resistance because of their higher crown positions, where they are exposed to higher solar radiation, greater wind speeds, lower humidity, and lower CO_2_ concentrations [@koike_leaf_2001; @kunert_revised_2017 **REFS-KAT**). Subcanopy trees subsequently tend to fare better specifically due to the benefits of a buffered environment [@pretzsch_drought_2018]. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems confer a potential advantage in terms of allowing greater access to water; however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. A final mechanism that could mediate tree size-related responses to drought is how hydraulic traits are distributed with respect to size [@meakem_role_2018; @liu_hydraulic_2019 [Roskilly et al. 2019](https://www.pnas.org/content/pnas/early/2019/06/14/1900734116.full.pdf)]. The pattern observed by @bennett_larger_2015 could be caused if the larger size classes were dominated by species less adapted to handle drought. Alternatively, larger size classes may be dominated by species that are inherently better adapted to greater biophysical challenges--as is the case in tropical moist forests of Panama, where larger size classes contain greater proportions of deciduous species [@condit_quantifying_2000; @meakem_role_2018]. Understanding the mechanisms driving the greater growth reductions of larger trees during drought will require sorting out the interactive effects of height, canopy position, soil water availability, and species' traits.
